---
title: "Biodiv_Port_Corsica_eDNA"
author: "Rachel Haderlé"
date: "2024-10-24"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##Beta-diversity
```{r}
## Load necessary libraries and data imports
library(readxl)
library(ggplot2)
library(tidyr)
library(dplyr)
library(ape)
library(adespatial)
library(vegan)
library(factoextra)
library("sf")
library("rnaturalearth")
library("rnaturalearthdata")
library("ggspatial")
library(tidyselect)
library(raster)
library(VennDiagram)
library(viridis)
library(reshape2)
library(fastDummies)
library(fishtree)
library(picante)
library(geiger)
loadNamespace("rfishbase") # Ensures functions from rfishbase are loaded
library(rtrees)
library(taxize)
library(cluster)
library(tibble)

## Beta diversity: Jaccard dissimilarity and turnover/nestedness calculations

# Import occurrence data
occurrences <- read.csv("occ_data.csv")

# Convert the data into a presence-absence matrix (1 = present, 0 = absent)
adne <- dcast(occurrences, Taxonomie ~ Station, value.var = "Taxonomie", fun.aggregate = length)

# Identify columns to modify (all except 'Taxonomie')
columns_to_modify <- setdiff(colnames(adne), "Taxonomie")

# Apply binary conversion to the selected columns (1 if present, 0 otherwise)
adne[columns_to_modify] <- lapply(adne[columns_to_modify], function(x) ifelse(x > 0, 1, x))

# Save the presence-absence matrix to a CSV file
write.csv(adne, "adne.csv", row.names = FALSE)

# Read the generated presence-absence matrix and transpose it for further analysis
jaccard <- read.csv("adne.csv", header = TRUE, row.names = 1)
jaccard <- t(jaccard)

# Convert all elements to numeric format
jaccard[] <- apply(jaccard, c(1, 2), as.numeric)

# Perform Jaccard dissimilarity analysis using binary (presence-absence) data
JAC <- beta.div.comp(jaccard, quant = FALSE, save.abc = FALSE)
JAC

# Retrieve the dissimilarity matrix (Jaccard distances)
JAC$D

## Heatmap of Jaccard dissimilarities

# Convert the Jaccard distance matrix into a dataframe for visualization
jaccard_matrix <- as.data.frame(as.table(as.matrix(JAC$D)))
colnames(jaccard_matrix) <- c("Station1", "Station2", "Distance")

# Replace "_Autumn" with "_PS" (for post-storm) in the station names
jaccard_matrix$Station1 <- gsub("_Autumn", "_PS", jaccard_matrix$Station1)
jaccard_matrix$Station2 <- gsub("_Autumn", "_PS", jaccard_matrix$Station2)

# Create a heatmap with a reversed purple color palette (0 = white, 1 = dark purple)
ggplot(jaccard_matrix, aes(x = Station1, y = Station2, fill = Distance)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(
    option = "magma",    # Use the 'magma' purple color palette
    limits = c(0, 1),    # Set the limits for the color scale
    name = "Dissimilarity\n(Jaccard Distance)",
    direction = -1       # Reverse the scale: low values are light, high values are dark
  ) +
  labs(x = NULL, y = NULL) +  # Remove axis labels
  theme_minimal() +            # Use a minimalistic theme
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),   # Rotate x-axis labels
    plot.title = element_text(hjust = 0.5),              # Center the plot title
    plot.subtitle = element_text(hjust = 0.5)            # Center the plot subtitle
  )

## Pairwise turnover/nestedness table

# Define station names (using consistent naming)
stations <- c("Calvi_Autumn", "Calvi_Spring", "L.Île.Rousse_Autumn", "L.Île.Rousse_Spring",
              "Saint.Florent_Autumn", "Saint.Florent_Spring", "STARESO_Autumn", "STARESO_Spring")

# Initialize an empty list to store results for each station pair
results <- list()

# Calculate Jaccard dissimilarity for each pair of stations
for (i in 1:(length(stations) - 1)) {
  for (j in (i + 1):length(stations)) {
    # Exclude rows and columns of stations not being compared
    jaccard_subset <- jaccard[-c(i, j), -c(i, j)]
    JAC <- beta.div.comp(jaccard_subset, quant = FALSE, save.abc = FALSE)
    
    # Save BDtotal and RichDif values for each pair of stations
    results[[paste(stations[i], stations[j], sep = ", ")]] <- c(JAC$part[4], JAC$part[5])
  }
}

# Create a results matrix to store turnover/nestedness results
results_matrix <- matrix(NA, nrow = length(stations), ncol = length(stations))
rownames(results_matrix) <- stations
colnames(results_matrix) <- stations

# Fill the matrix with BDtotal (top half) and RichDif (bottom half) values
for (i in 1:length(stations)) {
  for (j in 1:length(stations)) {
    if (i < j) {
      results_matrix[i, j] <- results[[paste(stations[i], stations[j], sep = ", ")]][1]  # BDtotal
    } else if (i > j) {
      results_matrix[i, j] <- results[[paste(stations[j], stations[i], sep = ", ")]][2]  # RichDif
    }
  }
}

# Convert the matrix to a dataframe for readability
results_df <- as.data.frame(results_matrix)

# Round values to 2 decimal places and export to a CSV file
write.csv(format(round(results_df, 2), nsmall = 2), file = "resultats_jaccard.csv", row.names = TRUE)

# Display the final table
print(results_df)

```

#################################################################################
##Biodiversity indicators (inspired by Dalongeville et al., 2022 https://doi.org/10.1111/1365-2664.14276)

Load biodiversity data and calculate presence-absence matrix
```{r}
library(reshape2)

# Load the occurrence data from a CSV file
occurrences <- read.csv("occ_data.csv")

# Convert the data into a presence-absence matrix (1 if present, 0 otherwise)
adne <- dcast(occurrences, Taxonomie ~ Station, value.var = "Taxonomie", fun.aggregate = length)

# Identify columns to modify (all except "Taxonomie")
columns_to_modify <- setdiff(colnames(adne), "Taxonomie")

# Convert the selected columns to binary (1 if present, 0 otherwise)
adne[columns_to_modify] <- lapply(adne[columns_to_modify], function(x) ifelse(x > 0, 1, x))

# Save the presence-absence matrix to a CSV file
write.csv(adne, "adne.csv", row.names = FALSE)
```


Load eDNA data and filter the species list
```{r}
## Load the eDNA data (matrix of species by sample)
adne <- read.csv("adne.csv", row.names = 1)

# Extract the list of species and samples
species <- rownames(adne)
samples <- colnames(adne)

# Load the species data again
adnei <- read.csv("adne.csv", header = TRUE)

# Filter species names that are too generic (e.g., "sp.") or only include family names
species_list <- adnei$Taxonomie
species_list <- adnei[!grepl("\\bsp\\.\\b", sapply(strsplit(as.character(species_list), " "), `[`, 2)) & 
                      sapply(strsplit(as.character(species_list), " "), length) > 1, ]

# Keep the filtered species names
species_list <- species_list$Taxonomie

# Check the filtered species names
head(species_list)

# Save the filtered species list
write.csv(species_list, file = "species_list.csv", row.names = FALSE)
```

Use OpenRefine for IUCN status and invasive species, then merge with traits data
```{r}
# Load species data from OpenRefine with IUCN status and invasive species information
species_list_open <- read.csv2("species_list_open.csv", header = TRUE)

# Rename columns for clarity
species_list_open <- species_list_open %>%
  rename(Species = x, IUCN_status = IUCN.conservation.status, invasive = invasive.to)

# Retrieve traits information from FishBase using the species list
traits_bruts <- rfishbase::species(species_list = species_list, fields = c("Species", "CommonLength", "Vulnerability", "DemersPelag", "Importance"))

# Get trophic level data from FishBase
trophic_data <- rfishbase::ecology(species_list, fields = c("Species", "FoodTroph"))
traits_bruts <- merge(traits_bruts, trophic_data[, c("Species", "FoodTroph")], by = "Species", all.x = TRUE)

# Merge FishBase traits with IUCN status data
traits_bruts <- traits_bruts %>%
  left_join(species_list_open, by = "Species")

# Save the merged traits and IUCN status data
write.csv(traits_bruts, file = "traits_bruts.csv", row.names = FALSE)

# Preview the first rows of the merged data
head(traits_bruts)
```

Combine species traits with eDNA data
```{r}
# Load the updated traits data
traits_bruts <- read.csv("traits_bruts.csv", header = TRUE)

# Load the original occurrence data
adnei <- read.csv("adne.csv", header = TRUE)

# Merge the occurrence data with species traits
adnei_with_traits <- merge(adnei, traits_bruts, by.x = "Taxonomie", by.y = "Species", all = TRUE)

# Save the merged data
write.csv(adnei_with_traits, file = "adnei_with_traits.csv", row.names = FALSE)
```


Perform biodiversity indicator calculations
```{r}
# Load the merged data with traits and replace NA values with 0 (except in 'Vulnerability')
adnei_with_traits <- read.csv("adnei_with_traits.csv", header = TRUE)

# Mark invasive species (binary: 1 for invasive, 0 otherwise)
adnei_with_traits <- adnei_with_traits %>%
  mutate(Invasive = ifelse(invasive == "1", 1, 0))
adnei_with_traits$Invasive[is.na(adnei_with_traits$Invasive)] <- 0

# Add columns for different IUCN categories
adnei_with_traits <- adnei_with_traits %>%
  mutate(
    IUCN_Red_List_Category_VU = ifelse(IUCN_status == "Vulnerable", 1, 0),
    IUCN_Red_List_Category_EN = ifelse(IUCN_status == "Endangered", 2, 0),
    IUCN_Red_List_Category_CR = ifelse(IUCN_status == "Critically Endangered", 3, 0)
  )

# Save the updated dataset
write.csv(adnei_with_traits, file = "adnei_with_traits_updated.csv", row.names = FALSE)

# Display the first few rows of the updated dataset
print(head(adnei_with_traits))
```

Create the indicators table
```{r}
# Create the result matrix for biodiversity indicators
indicators <- matrix(NA, ncol(adne), 9,
                     dimnames = list(colnames(adne),
                                     c("R", "PD", "PD/SR", "FD", "FD/R", "Vulner", "RedList", "TLmean", "Invasive")))
```

###Taxonomic richness
```{r}
indicators[,1] <- apply(adne, 2, sum)
```

###Functional Diversity
To calculate functional diversity (FD), we first retrieve species traits from FishBase, perform clustering using PCoA, and map them to MOTUs (Molecular Operational Taxonomic Units).
```{r}
# List of probable species from FishBase
fish_species <- species_list

# Retrieve trait data for each species from FishBase
all_traits_list <- rfishbase::species(species_list = fish_species, 
                                      fields = c("Species", "CommonLength", "Weight", "LongevityWild", 
                                                 "AnaCat", "MigratRef", "DepthRangeShallow", "DepthRangeDeep"))

# Convert the retrieved list to a data frame
traits_data <- as.data.frame(all_traits_list)

# Select the relevant columns for analysis
traits_subset <- traits_data %>%
  dplyr::select(Species, CommonLength, Weight, LongevityWild, AnaCat, 
                MigratRef, DepthRangeShallow, DepthRangeDeep)

# Select only the numeric columns for further processing
numeric_columns <- c("CommonLength", "Weight", "LongevityWild", "DepthRangeShallow", "DepthRangeDeep")

# Replace missing values in numeric columns with the median of each column
for (col in numeric_columns) {
  traits_subset[[col]][is.na(traits_subset[[col]])] <- median(traits_subset[[col]], na.rm = TRUE)
}

# Convert categorical variables to factors if needed
traits_subset$AnaCat <- as.factor(traits_subset$AnaCat)
traits_subset$MigratRef <- as.factor(traits_subset$MigratRef)

# Calculate trait dissimilarity using Gower's distance
gower_dist <- daisy(traits_subset[, -1], metric = "gower")

# Print the dissimilarity matrix (optional)
print(gower_dist)

# Perform Principal Coordinate Analysis (PCoA) on the dissimilarity matrix
pcoa_result <- pcoa(gower_dist)

# Extract the PCoA scores (coordinates of species in PCoA space)
pcoa_scores <- as.data.frame(pcoa_result$vectors)

# Set species names as row names
rownames(pcoa_scores) <- traits_subset$Species

# Select the first two PCoA axes for analysis
pcoa_scores <- pcoa_scores[, 1:2]

# Function to calculate functional diversity (variation in trait space)
functional_diversity <- function(x) {
  sqrt(sum((x - mean(x))^2))
}

# Calculate functional diversity for each species using the first two PCoA axes
diversity_scores <- apply(pcoa_scores, 1, functional_diversity)

# Round the functional diversity scores to 5 decimal places
diversity_scores <- round(diversity_scores, digits = 5)

# Create a data frame with species names and their corresponding functional diversity scores
diversity_results <- data.frame(Species = rownames(pcoa_scores), FunctionalDiversity = diversity_scores)

# Display the functional diversity results (optional)
print(diversity_results)

# Load the species occurrence data (adne dataset)
adnei <- read.csv("occ_data.csv", header = TRUE)

# Merge the occurrence data (adne) with the functional diversity results based on the species (Taxonomy)
traits_updated <- merge(adnei, diversity_results, by.x = "Taxonomie", by.y = "Species", all = TRUE)

# Rename the functional diversity column to "GF" (Functional Group)
colnames(traits_updated)[colnames(traits_updated) == "FunctionalDiversity"] <- "GF"

# Replace NA values in the "GF" column with 0 (indicating no functional group assigned)
traits_updated <- traits_updated %>%
  dplyr::mutate(GF = replace(GF, is.na(GF), 0))

# Save the updated data to a CSV file
write.csv(traits_updated, file="traits_updated.csv")
```

For each sample, the functional diversity is calculated based on the number of unique functional groups.
```{r}
# Calculate the functional diversity (FD) indicator for each sample
for (i in 1:ncol(adne)) { 
  # List the species present in the sample
  s_i <- adne %>%
    filter(adne[,i] == 1) %>%
    tibble::rownames_to_column(var="Sp") %>%
    pull(Sp)
  
  # Retrieve the functional groups (GF) of the species present
  fd_i <- as.factor(traits_updated[traits_updated$Taxonomie %in% s_i, "GF"])
  
  # Count the number of unique functional groups and store it in the indicators dataframe
  indicators[i, "FD"] <- length(unique(fd_i))
}

# (Optional) Normalize the FD indicator by species richness (R) for each sample
indicators[, "FD/R"] <- indicators[, "FD"]/indicators[, "R"]
```

###Invasive species
```{r}
# Loop to calculate the invasive species indicator for each sample
for (i in 1:nrow(indicators)) { 
  # List the species present in the sample
  s_i <- adne %>%
    filter(adne[,i] == 1) %>%
    tibble::rownames_to_column(var="Sp") %>%
    pull(Sp)
  
  # Sum the number of invasive species in the sample and store the result in the indicators dataframe
  indicators[i, "Invasive"] <- sum(adnei_with_traits[which(adnei_with_traits$Taxonomie %in% s_i), "Invasive"])
}
```

###TLmean
```{r}
# Calculate mean trophic level (TLmean) for each sample
for (i in 1:nrow(indicators)) {  # Loop through each sample
  # List species present in the sample
  s_i <- adne %>%
    filter(adne[, i] == 1) %>%
    tibble::rownames_to_column(var = "Sp") %>%
    pull(Sp)

  # Calculate mean trophic level
  indicators[i, "TLmean"] <- mean(adnei_with_traits[which(adnei_with_traits$Taxonomie %in% s_i), "FoodTroph"], na.rm = TRUE)
}

# Calculate mean and standard deviation for TLmean with formatted output
for (i in 1:nrow(indicators)) {
  # List species present in the sample
  s_i <- adne %>%
    filter(adne[, i] == 1) %>%
    tibble::rownames_to_column(var = "Sp") %>%
    pull(Sp)

  # Calculate mean and standard deviation for TLmean
  tl_values <- adnei_with_traits[which(adnei_with_traits$Taxonomie %in% s_i), "FoodTroph"]
  tl_mean <- mean(tl_values, na.rm = TRUE)
  tl_sd <- sd(tl_values, na.rm = TRUE)

  # Store formatted value with mean and standard deviation
  indicators[i, "TLmean"] <- sprintf("%.2f (± %.2f)", tl_mean, tl_sd)
}
```

###Vulnerability
```{r}
# Calculate vulnerability indicator for each sample
for (i in 1:nrow(indicators)) {  # Loop through each sample
  # List species present in the sample
  s_i <- adne %>%
    filter(adne[, i] == 1) %>%
    tibble::rownames_to_column(var = "Sp") %>%
    pull(Sp)

  # Calculate mean vulnerability for the species present
  indicators[i, "Vulner"] <- mean(adnei_with_traits[which(adnei_with_traits$Taxonomie %in% s_i), "Vulnerability"], na.rm = TRUE)
}

```

###IUCN Status
```{r}
# Calculate the number of species in each IUCN category for each sample
for (i in 1:nrow(indicators)) {  # Loop through each sample
  # List species present in the sample
  s_i <- adne %>%
    filter(adne[, i] == 1) %>%
    tibble::rownames_to_column(var = "Sp") %>%
    pull(Sp)

  # Count species in Vulnerable (VU), Endangered (EN), and Critically Endangered (CR) categories
  VU <- sum(adnei_with_traits[which(adnei_with_traits$Taxonomie %in% s_i), "IUCN_Red_List_Category_VU"], na.rm = TRUE)
  EN <- sum(adnei_with_traits[which(adnei_with_traits$Taxonomie %in% s_i), "IUCN_Red_List_Category_EN"], na.rm = TRUE)
  CR <- sum(adnei_with_traits[which(adnei_with_traits$Taxonomie %in% s_i), "IUCN_Red_List_Category_CR"], na.rm = TRUE)

  # Calculate total number of species in Red List categories
  indicators[i, "RedList"] <- (VU + EN + CR)
}

```

###Phylogenetic Diversity of Fish

```{r}

# Read the dataset
test <- read.csv("adne.csv", header = TRUE)

# Replace spaces with underscores in the species column
test$Taxonomie <- gsub(" ", "_", test$Taxonomie)

# Create a tibble for the species list
test_fish_list <- tibble::tibble(
  species = test$Taxonomie,
  genus = test$Genus,
  family = test$Family,
  order = test$Order
)

# Retrieve the phylogenetic tree for fish species
phy2 <- get_tree(sp_list = test_fish_list,
                 taxon = "fish",
                 scenario = "at_basal_node",
                 show_grafted = TRUE)

# Specify species to exclude from the tree
data_not_tree <- c("Mobula_mobular", "Sparidae")

# Read occurrence data
adne <- read.csv("adne.csv", header = TRUE, row.names = 1)
rownames(adne) <- gsub(" ", "_", rownames(adne))

# Filter out species not in the phylogenetic tree
adne2 <- adne[!rownames(adne) %in% data_not_tree, ]

# Check species names in adne2 against the phylogenetic tree
phy_labels <- phy2$tip.label
adne_labels <- rownames(adne2)

# Identify species present in adne2 but not in phy2
not_in_phy2 <- setdiff(adne_labels, phy_labels)
# Identify species present in phy2 but not in adne2
not_in_adne2 <- setdiff(phy_labels, adne_labels)

# Print species names in the phylogenetic tree
print(phy2$tip.label)

# Remove asterisks from species names in phy2
phy2$tip.label <- gsub("\\*", "", phy2$tip.label)

# Check modified species names
print(phy2$tip.label)

# Identify species to remove from the phylogenetic tree
missing_from_tree <- setdiff(adne_labels, phy_labels)
# Prune the tree to remove specified species
prunedTree <- drop.tip(phy2, missing_from_tree)

# Check the structure of the pruned tree
print(prunedTree)

# Filter adne2 to keep only species present in prunedTree
adne2_filtered <- adne2[rownames(adne2) %in% prunedTree$tip.label, ]

# Check the dimensions of adne2_filtered
print(dim(adne2_filtered))

# Transpose the ADNe matrix
adne2_filtered <- t(adne2_filtered)

# Calculate Faith's PD
pd.result <- picante::pd(adne2_filtered, prunedTree, include.root = TRUE)

# Check the results of PD calculation
print(pd.result)

# Store the PD values in the indicators dataframe
indicators[, "PD"] <- pd.result$PD

# Calculate the ratio of PD to species richness (SR)
indicators[, "PD/SR"] <- pd.result$PD / pd.result$SR

```

```{r}
write.csv(indicators, file="indicators_updated.csv")
```

################################################################################
##Jaccard and dissimilarity with functional group values.
```{r}
# Load the CSV file containing the occurrences
occurrences <- read.csv("traits_updated.csv")

# Load necessary libraries
library(reshape2)

# Create a presence-absence (PA) matrix
PA_matrix <- dcast(occurrences, Station ~ GF, fun.aggregate = length, value.var = "GF")

# Convert counts to presence-absence (1 if present, 0 otherwise)
PA_matrix[,-1] <- lapply(PA_matrix[,-1], function(x) ifelse(x > 0, 1, 0))

# Ensure that the columns are numeric
PA_matrix[-1] <- lapply(PA_matrix[-1], as.numeric)

# Define row names for the PA matrix using the Station column
row.names(PA_matrix) <- PA_matrix$Station

# Remove the Station column now that we have defined the row names
PA_matrix <- PA_matrix[,-1]

# Check the structure of the PA matrix
str(PA_matrix)

# Calculate beta diversity using Jaccard index
jacfd <- beta.div.comp(as.matrix(PA_matrix), quant = FALSE, save.abc = FALSE)

# Print the results
print(jacfd)

# List of stations to compare
stations <- c("Calvi_Autumn", "Calvi_Spring", "L.Île.Rousse_Autumn", "L.Île.Rousse_Spring",
              "Saint.Florent_Autumn", "Saint.Florent_Spring", "STARESO_Autumn", "STARESO_Spring")

# Create a list to store results for each pair of stations
results <- list()

# Calculate Jaccard index for each pair of stations
for (i in 1:(length(stations) - 1)) {
  for (j in (i + 1):length(stations)) {
    # Exclude rows and columns of non-compared stations
    jaccard_subset <- PA_matrix[-c(i, j), -c(i, j)]
    JAC <- beta.div.comp(as.matrix(jaccard_subset), quant = FALSE, save.abc = FALSE)
    
    # Store BDtotal and RichDif for each pair
    results[[paste(stations[i], stations[j], sep = ", ")]] <- c(JAC$part[4], JAC$part[5])  # JAC$part[4]: BDtotal, JAC$part[5]: RichDif
  }
}

# Create a cross-tabulation matrix for easy readability of results
results_matrix <- matrix(NA, nrow = length(stations), ncol = length(stations))
rownames(results_matrix) <- stations
colnames(results_matrix) <- stations

# Fill the matrix with the computed values
for (i in 1:length(stations)) {
  for (j in 1:length(stations)) {
    if (i < j) {
      results_matrix[i, j] <- results[[paste(stations[i], stations[j], sep = ", ")]][1]  # BDtotal
    } else if (i > j) {
      results_matrix[i, j] <- results[[paste(stations[j], stations[i], sep = ", ")]][2]  # RichDif
    }
  }
}

# Convert the results matrix to a data frame for better readability
results_df <- as.data.frame(results_matrix)

# Round values to 2 decimal places and export to a CSV file
write.csv(format(round(results_df, 2), nsmall = 2), file = "resultats_jaccard.csv", row.names = TRUE)

# Print the final results table
print(results_df)

statnam<- c("Calvi_Autumn", "Calvi_Spring", "L.Île.Rousse_Autumn", "L.Île.Rousse_Spring", 
              "Saint.Florent_Autumn", "Saint.Florent_Spring", "STARESO_Autumn", "STARESO_Spring")

# Perform PCA on the PA matrix
res.pca <- prcomp(PA_matrix)

# Visualize PCA results
RES.PCA <- fviz_pca_biplot(res.pca,
                            habillage = statnam,
                            labelsize = 3,
                            repel = TRUE,
                            legend.title = "Sample",
                            title = NULL) + 
           theme(text = element_text(size = 10))

# Display the PCA plot
RES.PCA
```

##Trophic levels - diagram
```{r}
# Assign a trophic category and convert NAs to "Unknown"
adnei_with_traits <- adnei_with_traits %>% 
  mutate(TrophicCategory = case_when(
    FoodTroph >= 2.0 & FoodTroph < 2.19 ~ "Herbivore",
    FoodTroph >= 2.2 & FoodTroph < 2.79 ~ "Omnivore",
    FoodTroph >= 2.8 ~ "Carnivore",
    TRUE ~ NA_character_  # Replace NAs with NA
  ))

# Restructure data to have stations in a single column
adnei_long <- adnei_with_traits %>% 
  pivot_longer(cols = 2:9, names_to = "Station", values_to = "Presence") %>% 
  filter(Presence == 1)  # Filter for presence only

# Add a column for the season
adnei_long <- adnei_long %>% 
  mutate(Season = ifelse(grepl("Spring", Station, ignore.case = TRUE), "Spring", "Autumn")) %>% 
  # Remove everything after the underscore in the Station column
  mutate(Station = sub("_.*$", "", Station)) %>% 
  # Create a combined variable for station and season without duplicates
  mutate(StationSeason = ifelse(Season == "Spring", 
                                paste(Station, "Spring"), 
                                paste(Station, "Autumn")))

# Ensure facets are in the desired order
adnei_long$StationSeason <- factor(adnei_long$StationSeason, 
                                    levels = c("Calvi Spring", "Calvi Autumn", 
                                               "L.Île.Rousse Spring", "L.Île.Rousse Autumn", 
                                               "Saint.Florent Spring", "Saint.Florent Autumn", 
                                               "STARESO Spring", "STARESO Autumn"))

# Summarize data by station and trophic category
trophic_summary <- adnei_long %>% 
  group_by(StationSeason, TrophicCategory) %>% 
  summarise(Count = n(), .groups = "drop") %>% 
  filter(!is.na(TrophicCategory)) %>%  # Remove NA categories
  group_by(StationSeason) %>% 
  mutate(RelativeFreq = Count / sum(Count))  # Calculate relative frequency

# Check summarized data
print(trophic_summary)

# Manually add TLmean values
trophic_summary <- trophic_summary %>% 
  mutate(TLmean = case_when(
    StationSeason == "Calvi Spring" ~ "3.18 (± 0.40)",
    StationSeason == "Calvi Autumn" ~ "3.38 (± 0.50)",
    StationSeason == "L.Île.Rousse Spring" ~ "3.33 (± 0.50)",
    StationSeason == "L.Île.Rousse Autumn" ~ "3.36 (± 0.53)",
    StationSeason == "Saint.Florent Spring" ~ "3.17 (± 0.52)",
    StationSeason == "Saint.Florent Autumn" ~ "3.05 (± 0.42)",
    StationSeason == "STARESO Spring" ~ "3.23 (± 0.37)",
    StationSeason == "STARESO Autumn" ~ "3.22 (± 0.46)",
    TRUE ~ NA_character_
  ))

# Replace dots with spaces and "Autumn" with "Post-storm"
trophic_summary <- trophic_summary %>% 
  mutate(StationSeason = gsub("\\.", " ", StationSeason),  # Remove dots
         StationSeason = gsub("Autumn", "Post-storm", StationSeason))  # Replace Autumn with Post-storm

# Define the order of stations to ensure Spring is before Post-storm
trophic_summary$StationSeason <- factor(trophic_summary$StationSeason, 
                                         levels = c("Calvi Spring", "L Île Rousse Spring", "Saint Florent Spring", "STARESO Spring", 
                                                    "Calvi Post-storm", "L Île Rousse Post-storm", "Saint Florent Post-storm", "STARESO Post-storm"))

# Create a pie chart with TLmean (sd) values at the center
ggplot(trophic_summary, aes(x = "", y = RelativeFreq, fill = TrophicCategory)) + 
  geom_bar(stat = "identity", width = 1) + 
  coord_polar("y") + 
  facet_wrap(~ StationSeason, nrow = 2, ncol = 4) +  # Spring on the first row, Post-storm on the second
  scale_fill_manual(values = c("Herbivore" = "#FF9999", "Omnivore" = "#FFCC99", "Carnivore" = "#99CCFF")) +  # Pastel color palette
  labs(fill = "Trophic Category") + 
  theme_void() +  # Display TLmean (sd) at the center of each circle
  geom_text(aes(label = TLmean), x = 0, y = 0, color = "black", size = 3, inherit.aes = FALSE)  

```
